\section{Proof Sketches and Deferred Proofs}

This appendix records the short proofs whose main ideas already appear in the
body, together with the claim-strength boundaries that matter for review.

\subsection{Uniform perturbation calculus}

The proof of \Cref{prop:replace-delete-add} is intentionally elementary. Its
value is not technical depth but category control: replace-one is an operation
composition, whereas LOO is an operation-plus-evaluation convention. The
triangle-inequality argument therefore applies directly to delete/add versus
replace, but not to LOO without additional translation lemmas.

\subsection[Deterministic k-NN perturbation criterion]{Deterministic $k$-NN perturbation criterion}

The proof of \Cref{prop:margin-crossing} reduces the classifier to a signed vote
threshold. All geometric complexity is compressed into the ordered top-$k$
neighbor list $N_k(S,x)$. Once that list is fixed, the classifier is just the
sign of $M_k(S,x)$ with the non-positive side resolved toward label $0$. The
only way a perturbation changes the prediction is therefore to move the signed
margin across zero.

\subsection[Two-point 1-NN witness]{Two-point $1$-NN witness}

The witness of \Cref{prop:two-point-separation} is fully explicit and
hand-checkable. No exhaustive search is needed for the existence claim itself.
What the repository's search artifacts contribute is different: they show that
within the enumerated search spaces, no one-vertex witness appears and the
two-vertex pattern is the first observed regime where the separation occurs.
That latter statement remains certificate-level evidence, not a proof.
